Zero derivative and local risk aversion

### 2021-04-01 ZeroDerivative.tex
Add files via upload

@@

@@

@@

@@

\date{}
\author{Danil Fedchenko}

@@

	Consider a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Denote $\pi(t)$ a risk premium, an expected-utility maximizer with Bernoulli utility $u(x)$ is willing to pay to avoid risk $t \varepsilon$ (i.e. $u(x-\pi(t)) = \int u(x+t\varepsilon)dF(\varepsilon)$). As \cite{segal1990first} stated, $\pi'(0) = 0$. This is true, however, only if $u'(x) \neq 0$. In this note I show that an attitude toward risk of an individual with $u'(x) = 0$ in some cases is similar to being what \cite{segal1990first} called of order 1. Which in particular implies that such an individual may find optimal to buy full insurance even for a small positive risk in the neighborhood of the points where $u'(x) = 0$.

@@

	Consider an individual with expected utility preferences represented by strictly increasing, continuous Bernoulli utility function $u(x)$. Suppose this individual for some $t \in \mathbb{R}$ faces a lottery $t \eps$ represented by a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Define a risk premium $\pi(t)$ implicitly as $x-\pi(t) \sim x+t\eps$ or equivalently $u(x - \pi(t)) = \int u(x + t \eps)\ dF(\eps)$ (from now on whenever I write $\int$, I assume the integral exists). It is easy to show that as long as $u'(x) \neq 0$, $\pi'(0) = 0$. Indeed, 

@@

	\pi(t) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right),\ u^{-1} \text{ exists since }u(x)\text{ is strictly increasing and continuous}\\
	\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon) = \\

@@

	\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} &= -\frac{u'(x)}{u'(x)} \expect[\eps] = -\expect[\eps] = 0, \text{ if } u'(x) \neq 0

@@

This in particular implies that for small enough $t$ $$\pi(t) \approx -\frac{t^2 \expect[\eps^2]}{2}\frac{u''(x)}{u'(x)}$$ and hence it approaches zero faster than $t$ which, in turn, implies that for small $t$ an individual is almost risk neutral. This assertion, however, heavily relies upon the assumption $u'(x) \neq 0$ which is known sometimes to be violated even for strictly increasing functions ($u(x) = x^3$ at $x=0$ is probably the most famous example). 
In this paper I partially characterize $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0}$ at points where $u'(x) = 0$, and show that it is not equal to zero in general. Applying this to the optimal insurance problem, as \cite{segal1990first} show, means that an expected utility maximizer may find optimal to buy full insurance even for small risk and even if insurance includes positive loading. I illustrate this result by deriving a slope of indifference curves (provided that it is non zero and finite) of the individual near the certainty line on the state-contingent  \cite{yaari1978some}-type diagram. It turns out that the slope differs from the case where $u'(x) \neq 0$.
\section{Order of Risk Aversion}
\cite{segal1990first} define an attitude toward risk of individual being of order 1 if for \textit{every} non-degenerate random variable $\eps$ such that $\expect[\eps] = 0$, $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} \neq 0$, and of order 2 if $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0^+} = 0$. 
In the next proposition I derive $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ (assuming that it exists, is non-zero and finite) at the points where for some $k\ge 1$ first $k-1$ derivatives of the Bernoulli utility function $u(x)$ exist and equal to zero.

@@

		If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$, and there exists $u^{(k)}(x^*) \neq 0$. Then if there exists finite and non-zero $\expect[\eps^k]$, and at $x^*$ there exists finite, non-zero $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}$ then :

@@

			\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^k = (-1)^k\ \expect[\varepsilon^k]

@@

	Provided that \begin{align}\label{limsup}
	\forall\ i = \{2,\dots, k-1\}:\ \bigg|\lim_{t\to0} \inf \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty, \bigg|\lim_{t\to0} \sup \frac{\partial^{(i)} \pi(t)}{\partial t^{(i)}}\bigg| < \infty
\end{align} for $k>2$.

@@

@@

		\pi(t) &= x^* - u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\\
		\frac{\partial \pi(t)}{\partial t} &= -\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon) = \\
		&=-\frac{\int u'(x^*+t \varepsilon)\varepsilon dF(\varepsilon)}{u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right)}\ ,\ \ \  \text{if }u'\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \neq 0

@@

As $t \to 0$ since $u'(x^*) = 0$, both numerator and denominator approach zero.
Using the L'Hospital rule we get:

@@

		\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u''(x^*+t \varepsilon) \varepsilon^2 dF(\varepsilon)}{u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \underbrace{\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right) \right)'}_{-\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}}}

@@

If $u''(x^*) \neq 0$ (i.e. $k=2$) then $\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \right)^2 = \expect[\varepsilon^2]$. If $k>2$ then one should apply L'Hospital rule once again. That is,

@@

	\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} = \lim_{t \to 0} -\frac{\int u'''(x^*+t \varepsilon) \varepsilon^3 dF(\varepsilon)}{u'''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0}\right)^2 - u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} }

@@

Since $u''(x^*) = 0$ and \eqref{limsup} implies  $$\lim_{t \to 0} u''\left(u^{-1}\left( \int u(x^* + t \varepsilon) dF(\varepsilon) \right)\right) \frac{\partial^2\pi(t)}{\partial t^2} = 0,$$ if $k=3$ then we have $$\left( \frac{\partial \pi(t)}{\partial t} \bigg|_{t=0}\right)^3 = -\expect[\eps^3] $$ which is what we sought to prove. If $k>3$ we should proceed in the same way, i.e. by applying L'Hospital rule $k-3$ times we get what we wanted to prove.

@@

Although, the proposition above does not tell us whether $\frac{\partial \pi(t)}{\partial t} \bigg|_{t = 0} \neq 0$, so we cannot conclude that the individual's attitude toward risk is necessary of order 1 (according to \cite{segal1990first}). We will see in Section 4 that, in general neither $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$ nor $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} = 0$, so the individual's attitude toward risk is neither of order 1 nor of order 2. Nevertheless, if it happens to be $\expect[\eps^k] \neq 0$ and $\frac{\partial \pi(t)}{\partial t}\bigg|_{t=0} \neq 0$, optimal behavior of such an individual regarding the choice of insurance is similar to that of the one whose attitude toward risk is of order 1.
\section{Optimal Insurance}
Discussing the optimal insurance problem consider the lottery $x^*+t \eps $ where $\expect[\eps] > 0$. Then, the following proposition is an analogue to the Proposition 1 of \cite{segal1990first}.

@@

	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, there exist non-zero left and right derivatives $u^{(k)}_{+}(x^*)$ and $u^{(k)}_{-}(x^*)$, and $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus\{x^*\}$ for some $\delta>0$. If $\eps$ possesses all finite moments, and in particular, there exist $\expect[ \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$ and \begin{align}\label{condition}
		u^{(k)}_{+}(x^*) \cdot \expect[ \eps^k | \eps\ge0]\prob(\eps\ge0) + u^{(k)}_{-}(x^*) \cdot \expect[ \eps^k | \eps<0]\prob(\eps<0) < 0,

@@

		\int u(x^*+t\eps)dF(\eps) = \int \left( u(x^*) +  \bigg[u^{(k)}_{+}(x^*)\mathbb{I}[\eps\ge0]  + u^{(k)}_{-}(x^*)\mathbb{I}[\eps<0] \bigg] \frac{t^k \eps^k }{k!} +o_p(t^k \eps^k)\right)dF(\eps) = \\
	= u(x^*) + \frac{(u^{(k)}_{+}(x^*) \expect[\eps^k|\eps\ge 0]\prob(\eps\ge0)+u^{(k)}_{-}(x^*) \expect[\eps^k|\eps< 0]\prob(\eps<0))t^k}{k!} + \expect[o_p(t^k \eps^k)] < u(x^*)

@@

where the first equality follows from Taylor formula and the fact that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, and the last inequality follows from \eqref{condition}. Note that for the $k$ times differentiable $u(x)$, \eqref{condition} becomes $u^{(k)}(x^*)\cdot\expect[\eps^k] < 0$

@@

The result above means that the expected-utility maximizer under some conditions exhibits local risk aversion at the vicinity of the points where his strictly increasing Bernoulli utility function fails to have positive derivatives up to the order $k$. In other words, this individual prefers a certain outcome to the lottery with small positive expected value. This result stands in contrast to the well-known result that the expected-utility maximizer with increasing Bernoulli utility function $u(x)$ is locally risk neutral as long as $u'(x) > 0$. That is, for $t$ small enough he would strictly prefer the lottery with positive mean to the certain outcome.

Let me illustrate the proposition \ref{LRA} by deriving slopes of the individual's indifference curves at a certainty line on the state-contingent diagram. 
Consider an individual whose expected utility preferences represented by strictly increasing and continuous Bernoulli utility function $u(x)$. Suppose there are two states of nature, the first happens with probability $p$ and the second with probability $1-p$. At state $i \in \{1, 2\}$ the individual gets $x_i$. Then the following proposition holds:
\begin{proposition}\label{slope}
	If $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some small $\delta>0$. If there exist left and right derivatives $u_{+}^{(k)}(x^*) \neq 0$, $u_{-}^{(k)}(x^*) \neq 0$ such that \footnote{$$sign(x) = \begin{cases}
			-1, x <0\\
			0, x=0\\
			1, x>0
		\end{cases}$$}$\text{sign}(u_{+}^{(k)}(x^*) \cdot u_{-}^{(k)}(x^*)) = (-1)^{k+1}$

	Then consider an indifference curve $x_2(x_1)$ implicitly given by 
	\begin{align}\label{ind}
		\{ (x_1, x_2): pu(x_1)+(1-p)u(x_2) = u(x^*) \}
	\end{align}
then if there exist finite, non-zero $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$,  $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0}$:
	\begin{align}
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{+}(x^*)}{u^{(k)}_{-}(x^*)} \bigg|  }\\
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{-}(x^*)}{u^{(k)}_{+}(x^*)} \bigg|  }
	\end{align}
Provided that 
\begin{align}\label{der}
\forall\ i = \{1, \dots, k-1\}:\ \bigg| \lim_{x_1 \to x^* \pm 0} \sup \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty, \bigg| \lim_{x_1 \to x^* \pm 0} \inf \frac{d^{(i)}x_2(x_1)}{dx_1^{(i)}}\bigg| < \infty
\end{align}
\end{proposition}
\begin{proof}
	From \eqref{ind}, implicit function theorem and the fact that $u'(x) > 0$ for $x \in (x^*-\delta, x^*+\delta)\setminus \{x^*\}$ it follows that
	\begin{align*}
		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = \lim_{x_1 \to x^*+0} -\frac{p}{1-p} \frac{u'(x_1)}{u'(x_2(x_1))}
	\end{align*}
Since $x_2(x^*) = x^*$ and $u'(x^*) = 0$ both numerator and denominator approach zero as $x_1 \to x^*+0$. Then if $	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$ exists we can use L'Hospital rule, i.e.
\begin{align*}
		\lim_{x_1 \to x^*+0} \frac{dx_2}{dx_1} = -\frac{p}{1-p} \frac{u''_{+}(x_1)}{u''_{-}(x_2(x_1)) \left(\frac{dx_2}{dx_1}\right)\bigg|_{x^*+0}}
	\end{align*}
We have left second derivative in the denominator since from the fact that $u(x)$ is strictly increasing it follows that the indifference curve should be downward slopping. Meaning that if $x_1$ approaches $x^*$ from above, $x_2(x_1)$ approaches $x^*$ from below.

If $k = 2$ and $u''_{+}(x^*) \cdot u''_{-}(x^*) < 0$ then $$	\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt{\frac{p}{1-p} \bigg|  \frac{u''_{+}(x^*)}{u''_{-}(x^*)} \bigg|  }$$
Minus sign again follows from downward slopping indifference curve. Left derivative can be found in exactly the same way.

>>>>

\usepackage[affil-it]{authblk}

@@

\renewcommand*{\Affilfont}{\normalfont}

@@

\newcommand{\posder}{u^{(k)}_{+}}
\newcommand{\negder}{u^{(k)}_{-}}

@@

\newtheorem{observation}[theorem]{Observation}

@@

\author{Danil Fedchenko\thanks{I thank Eddie Dekel for help and suggestions. Any remaining errors are mine. }}
\affil{Northwestern University}
\date{\today}

@@

	Consider a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Denote $\pi(t, x)$ a risk premium, an expected-utility maximizer with Bernoulli utility $u(x)$ is willing to pay to avoid risk $t \varepsilon$ (i.e. $u(x-\pi(t, x)) = \int u(x+t\varepsilon)dF(\varepsilon)$). As \cite{segal1990first} stated $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} = 0$ for all $x$, provided that $u'(x) \neq 0$. In this note I show that an attitude toward risk of an individual with $u'(x) = 0$ typically is similar to what \cite{segal1990first} called of order 1. This in particular implies that such an individual may find optimal to buy full insurance even for a small positive risk in the neighborhood of the points where $u'(x) = 0$.

@@

	Consider an individual with expected utility preferences represented by strictly increasing, continuous Bernoulli utility function $u(x)$. Suppose this individual for some $t \in \mathbb{R}$ faces a lottery $t \eps$ represented by a random variable $\varepsilon$ such that $\expect[\varepsilon] = 0$. Define a risk premium $\pi(t, x)$ implicitly as $x-\pi(t, x) \sim x+t\eps$ or equivalently $u(x - \pi(t, x)) = \int u(x + t \eps)\ dF(\eps)$ (from now on whenever I write $\int$, I assume the integral exists). It is easy to show that as long as $u'(x) \neq 0$, $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} = 0$. Indeed, 

@@

	\pi(t, x) &= x - u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \footnotemark\\
	\frac{\partial \pi(t, x)}{\partial t} &= -\left(u^{-1}\left( \int u(x + t \varepsilon) dF(\varepsilon) \right) \right)' \int u'(x+t \varepsilon)\varepsilon dF(\varepsilon) = \\

@@

	\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0} &= -\frac{u'(x)}{u'(x)} \expect[\eps] = -\expect[\eps] = 0, \text{ if } u'(x) \neq 0

@@

\footnotetext{$u^{-1} \text{ exists since }u(x)\text{ is strictly increasing and continuous}$}
Using this and continuing similarly one can show that for small enough $t$ $$\pi(t, x) \approx -\frac{t^2 \expect[\eps^2]}{2}\frac{u''(x)}{u'(x)}$$ and hence it approaches zero faster than $t$ which, in turn, implies that for small $t$ an individual is almost risk neutral. This assertion, however, relies upon the assumption $u'(x) \neq 0$ which may not only be violated for strictly increasing functions at some point ($u(x) = x^3$ at $x=0$ is probably the most famous example) but even more so, there exist strictly increasing and continuous functions whose derivative equals zero almost everywhere.

In this paper, in Section \ref{order},  I derive $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0}$ at points where $u'(x) = 0$, and show that it is not equal to zero in general\footnote{One might mistakenly conjecture  that if $\frac{\partial \pi(t, x_n)}{\partial t}\bigg|_{t=0}=0$ for some $x_n \underset{n \to \infty}{\to}x$, then $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{t=0}=0$, but our results show such a conjecture is typically false. Which means that $\frac{\partial \pi(t, x)}{\partial t}$ typically is not continuous at $(0, 0)$}. In Section \ref{insurance}, I apply this to the optimal insurance problem and show that an expected utility maximizer whose Bernoulli utility has zero derivative at some point may exhibit local risk aversion. That is, such an individual may find optimal to buy full insurance even for small risk and even if insurance includes positive loading. I illustrate this result by deriving a slope of indifference curves of the individual near the certainty line on the state-contingent  \cite{yaari1978some}-type diagram where we see that the slope differs from the case where $u'(x) \neq 0$. In Section \ref{Bsingular}, I do the same exercise considering an individual whose Bernoulli utility function is strictly increasing, continuous but has zero derivative almost everywhere. Finally, in Section \ref{Examples}, I consider several concrete examples which illustrate general results derived in the previous sections. 
\section{Order of Risk Aversion}\label{order}
\cite{segal1990first} define an attitude toward risk of individual being of order 1 if for every non-degenerate random variable $\eps$ such that $\expect[\eps] = 0$, $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0^+} \neq 0$ for all $x$, and of order 2 if $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0^+} = 0$ for all $x$. 
In the next proposition I derive $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0}$ (assuming that it exists) at a point $x^*$ where for some $k\ge 1$ the first $k-1$ derivatives of the Bernoulli utility function $u(x)$ exist and equal zero. Hereafter I assume that $\pi(t, x)$ is continuous in $t$ at $x^*$.

@@

		Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, while $u^{(k)}(x^*) \neq 0$. Then if $\expect|\eps^k| < \infty$ and at $x^*$ there exists $\frac{\partial \pi(t, x)}{\partial t} \bigg|_{t = 0}$ :

@@

			\left( \frac{\partial \pi(t, x)}{\partial t} \Bigg|_{\substack{t = 0\\x = x^*}} \right)^k = (-1)^k\ \expect[\varepsilon^k]

@@

@@

	By definition of $\pi(t, x)$

@@

		u(x-\pi(t, x)) &= \int u(x+t\eps)dF(\eps)

@@

expanding this in the neighborhood of $t=0$ for $x=x^*$ and using $\pi(0, x^*) = 0, u^{(i)}(x^*) = 0\ \forall\ i\in\{1, \dots, k-1\}$, we get
\begin{align}\label{deriva}
	u(x^*) + (-1)^k \frac{u^{(k)}(x^*)}{k!}\pi(t, x^*)^k + o(\pi(t, x^*)^k) &= \int \left(u(x^*) + \frac{u^{(k)}(x^*)t^k\eps^k}{k!}+o(t^k\eps^k) \right)dF(\eps) \nonumber\\
(-1)^k \frac{u^{(k)}(x^*)}{k!}\pi(t, x)^k + \pi(t, x^*)^ko(1) &= \int \left(\frac{u^{(k)}(x^*)t^k\eps^k}{k!}+t^k\eps^ko_{\eps}(1) \right)dF(\eps)
where all $o(\cdot)$ are defined as $t \to 0$.
Since for $t$ small enough $|\eps^ko_{\eps}(1)| < |\eps^k|$ and $\expect|\eps^k|<\infty$, by the dominated convergence theorem $\lim_{t\to0}\expect[\eps^ko_{\eps}(1)] = \expect[\lim_{t\to0}\eps^ko_{\eps}(1)] = 0$, i.e. $\expect[\eps^ko_{\eps}(1)] = o(1), t\to0$.
Then \eqref{deriva} implies
\begin{align}\label{deriv}
	\left(\frac{\pi(t, x^*)}{t}\right)^k = \frac{\left( \frac{u^{(k)}(x^*)\expect[\eps^k]}{k!} + o(1) \right)}{\left( (-1)^k \frac{u^{(k)}(x^*)}{k!} + o(1)) \right)}
Since

@@

	\frac{\partial \pi(t, x^*)}{\partial t}\bigg|_{t = 0} = \lim_{t\to0}\frac{\pi(t, x^*)}{t}

@@

from \eqref{deriv} it follows that

@@

	\left( \frac{\partial \pi(t, x)}{\partial t} \bigg|_{\substack{t = 0\\x = x^*}} \right)^k = (-1)^k\ \expect[\varepsilon^k]

@@

@@

Thus, if $u'(x) = 0$ the individual's attitude toward risk is in general neither of order 1 nor of order 2, according to \cite{segal1990first} (e.g. for $k=3$ it could be the case $\expect[\eps^3] = 0$) Nevertheless, if it happens to be $\expect[\eps^k] \neq 0$ (where $k$ is defined as in the proposition above) and $\frac{\partial \pi(t, x)}{\partial t}\bigg|_{\substack{t=0\\x = x^*}} \neq 0$, optimal behavior of such an individual regarding the choice of insurance in a neighborhood of $x^*$ is similar to that of the one whose attitude toward risk is of order 1.
\section{Optimal Insurance}\label{insurance}
Discussing the optimal insurance problem, for $t \ge 0$ consider the lottery $x^*+t \eps $ where \textbf{$\expect[\eps] > 0$}. Then, the following proposition is an analogue to Proposition 1 of \cite{segal1990first}.

@@

	Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus\{x^*\}$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$. In addition, there exist finite left and right derivatives $u^{(k)}_{+}(x^*)$ and $u^{(k)}_{-}(x^*)$ such that at least one of these derivatives is non-zero. If there exist finite $\expect| \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$ and \begin{align}\label{condition}
		u^{(k)}_{+}(x^*) \cdot \expect[ \eps^k | \eps\ge0]\ \prob(\eps\ge0) + u^{(k)}_{-}(x^*) \cdot \expect[ \eps^k | \eps<0]\ \prob(\eps<0) < 0,

@@

		\int u(x^*+t\eps)dF(\eps) = \int \left( u(x^*) +  \bigg[u^{(k)}_{+}(x^*)\mathbb{I}[\eps\ge0]  + u^{(k)}_{-}(x^*)\mathbb{I}[\eps<0] \bigg] \frac{t^k \eps^k }{k!} +o(t^k \eps^k)\right)dF(\eps) = \\
	= u(x^*) + \frac{(u^{(k)}_{+}(x^*) \expect[\eps^k|\eps\ge 0]\prob(\eps\ge0)+u^{(k)}_{-}(x^*) \expect[\eps^k|\eps< 0]\prob(\eps<0))t^k}{k!} + t^k\expect[\eps^k o_{\eps}(1)] < u(x^*)

@@

where the first equality follows from Taylor formula and the fact that $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$, and the last inequality follows from \eqref{condition} and that $\lim_{t\to0}\expect[\eps^ko_{\eps}(1)]=0$, as in the proof of Proposition \ref{derivative} (here $\expect|\eps^k| < \infty$, since there exist finite $\expect[ \eps^k |\eps\ge0]$, $\expect[ \eps^k |\eps<0]$).

@@

Note that for a $k$ times differentiable $u(x)$, \eqref{condition} becomes $u^{(k)}(x^*)\cdot\expect[\eps^k] < 0$, meaning that if $k$ is the order of first non-zero derivative of $u(x)$ then its sign times the sign of the $k$-th moment determines if the person is locally risk loving or risk averse. 
In other words, the result above means that the expected-utility maximizer under some conditions exhibits local risk aversion (or risk loving) at the vicinity of the points where his strictly increasing Bernoulli utility function fails to have positive derivatives up to the order $k$. Consequently, this individual (in the case of risk-aversion) prefers a certain outcome to the lottery with a small positive expected value. This result stands in contrast to the well-known result that the expected-utility maximizer with increasing Bernoulli utility function $u(x)$ is locally risk neutral as long as $u'(x) > 0$. That is, for $t$ small enough he would strictly prefer the lottery with a positive mean to the certain outcome.

I illustrate Proposition \ref{LRA} by deriving slopes of the individual's indifference curves at a certainty line on the state-contingent diagram. 
Consider an individual whose expected utility preferences are represented by strictly increasing and continuous Bernoulli utility function $u(x)$. Suppose there are two states of nature, the first happens with probability $p$ and the second with probability $1-p$. At state $i \in \{1, 2\}$ the individual gets $x_i$. Then the following proposition holds:
\begin{proposition}\label{slope}
	Suppose $\exists\ x^* \in \mathbb{R}$ and $k \ge 2$ such that $u(x)$ is $k$ times differentiable on $(x^*-\delta, x^*+\delta)\setminus \{x^*\}$ for some $\delta>0$, and $\forall\ i = \{1, \dots, k-1  \}$: $u^{(i)}(x^*) = 0$. In addition, there exist left and right derivatives: $u_{-}^{(k)}(x^*)$, $u_{+}^{(k)}(x^*)$ such that  at least one of these derivatives is non-zero.

	Then consider an indifference curve $x_2(x_1)$ implicitly given by 
	\begin{align}\label{ind}
		\{ (x_1, x_2): pu(x_1)+(1-p)u(x_2) = u(x^*) \}
	\end{align}
then if there exist $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0}$,  $\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0}$ then
	\begin{align}
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*+0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{+}(x^*)}{u^{(k)}_{-}(x^*)} \bigg|  }\\
		\frac{d x_2}{d x_1} \bigg|_{x_1=x^*-0} = -\sqrt[k]{\frac{p}{1-p} \bigg|  \frac{u^{(k)}_{-}(x^*)}{u^{(k)}_{+}(x^*)} \bigg|  }
	\end{align}
\end{proposition}
In the proof of this proposition we will need the following lemma
\begin{lemma}\label{lemma}
	Under conditions of Proposition \ref{slope}:
	$$\footnote{$sign(x) = \begin{cases} -1, &x < 0\\ 0, &x = 0\\ 1, &x > 0 \end{cases}$}sign\left(\frac{\posder(x^*)}{\negder(x^*)} \right) \neq (-1)^k$$
	where we assume that ($sign(-\infty) = -1, sign(+\infty) = 1$)
\end{lemma}
\begin{proof}
	Suppose $k=2$, without loss of generality assume $x^* = 0$ and $u(x^*) = 0$. Then $u'(0) = 0$. Since $u$ is twicely differentiable to the left and to the right of zero, there should exist 
	\begin{align*}
		u''_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u'(\varepsilon)}{\varepsilon},\  u''_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u'(\varepsilon)}{\varepsilon}
		\end{align*}
	By assumption, $u$ is strictly increasing and differentiable in a neighborhood of $x^*$ which means that $u'(x) \ge 0$. Hence \begin{align*}
		u''_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u'(\varepsilon)}{\varepsilon} \le 0\\
		u''_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u'(\varepsilon)}{\varepsilon} \ge 0
	\end{align*}
and consequently $sign\left(\frac{u''_{+}(x^*)}{u''_{-}(x^*)} \right) \neq 1$.

Suppose $k = 3$ then $u'(0) = u''(0) = 0$ and as we saw for the case of $k=1$, 
 \begin{align*}
	u''_{-}(0) \le 0,\ u''_{+}(0) \ge 0
\end{align*}
then \begin{align*}
	u^{(3)}_{-}(0) = \lim_{\varepsilon \to 0^-} \frac{u''(\varepsilon)}{\varepsilon} \ge 0\\
	u^{(3)}_{+}(0) = \lim_{\varepsilon \to 0^+} \frac{u''(\varepsilon)}{\varepsilon} \ge 0
\end{align*}
meaning that $sign\left(\frac{u^{(3)}_{+}(x^*)}{u^{(3)}_{-}(x^*)} \right) \neq -1$. Proceeding in the same way, we can prove the result for an arbitrary $k$.
\end{proof}
\begin{proof} of Proposition \ref{slope}

	Consider small $\epsilon > 0$, then
	\begin{align*}
		u(x^* + \epsilon) = u(x^*) + \frac{u^{(k)}_{+}(x^*) \epsilon^k}{k!} + o(\epsilon^k)\\
		u(x^* + \epsilon)= u(x^*)+\epsilon^k\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right)
	\end{align*}
where $o(\cdot)$ is defined as $\epsilon \to 0$.
Similarly, for small $\omega <0$
\begin{align*}
	u(x^* + \omega) = u(x^*) + \frac{u^{(k)}_{-}(x^*) \omega^k}{k!} + o(\omega^k)\\
	u(x^* + \omega) = u(x^*) + \omega^k\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right)
\end{align*}
where $o(\cdot)$ is defined as $\omega \to 0$.
Consider $x_1 = x^*+\epsilon$, then since $u(x)$ is strictly increasing implies that $x_2 = x^* + \omega(\epsilon)$. Then,
\begin{align*}
	\frac{dx_2}{dx_1}\bigg|_{x_1=x^*+0} = \lim_{\epsilon \to 0^+} \frac{\omega(\epsilon)}{\epsilon}
\end{align*}
and \eqref{ind} implies:
\begin{align*}
		p\epsilon^k\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right) + (1-p)\omega^k(\epsilon)\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right) = 0\\
		\left(\frac{\omega(\epsilon)}{\epsilon}\right)^k = -\frac{p}{1-p} \frac{\left( \frac{u^{(k)}_{+}(x^*)}{k!} + o(1) \right)}{\left( \frac{u^{(k)}_{-}(x^*)}{k!} + o(1) \right)}
\end{align*}
By Lemma \ref{lemma} $sign\left(\frac{\posder(x^*)}{\negder(x^*)} \right) \neq (-1)^k$ then 
\begin{align*}
		\frac{dx_2}{dx_1}\bigg|_{x_1=x^*+0} = \lim_{\epsilon \to 0^+}\frac{\omega(\epsilon)}{\epsilon} = -\sqrt[k]{\frac{p}{1-p} \bigg|\frac{\posder(x^*)}{\negder(x^*)} \bigg| }
\end{align*}
$\frac{dx_2}{dx_1}\bigg|_{x_1=x^*-0}$ can be found in exactly the same way.
\end{proof}
In contrast to the result above it is known that in case of differentiable Bernoulli utility, the slope of an expected-utility maximizer's indifference curve at certainty points where $u'(x) \neq 0$ is equal to $-\frac{p}{1-p}$. If this individual is offered to buy insurance which would cost $\gamma$ for every insured unit and pay $\$\ 1$ in case of the bad event occurs, then the individual's budget constraint would have a slope $-\frac{\gamma}{1-\gamma}$. This in particular explains why such an expected-utility maximizer never buys full insurance unless it is actuarially fair (i.e. $\gamma = p$). At the same time, as we see from Proposition \ref{slope} an individual with $k$ times differentiable Bernoulli utility with the first $k-1$ derivatives equal to zero at some $x$ has slope of the indifference curve at $(x, x)$ given by $-\sqrt[k]{\frac{p}{1-p}}$. So depending on $p$ there may exist a range of positive loading $\gamma > p$ for which the expected utility maximizer would find it optimal to buy full insurance. In terms of Proposition \ref{LRA}, this range corresponds to the range of lotteries with positive expectation satisfying condition \eqref{condition}. Section \ref{Examples} illustrates Propositions \ref{LRA} and \ref{slope} by concrete examples.
